Appendix 2 - Fitting proposed path models using the C-statistic

### 2016-01-26 ch2_path_appendix2.Rmd
Edit to incorporat table of contents.

---
title: "Appendix 2 - Fitting proposed path models"
author: "Whalen Dillon"
date: "December 17, 2014"
output: html_document
---
----
#### This appendix details steps 1-4 of calculating the *C*-statistic for fitting a path model in a generalized multilevel context (Shipley 2009).

I go through these steps for the path models in Fig. 7 and Fig. 8 (included here)  to produce the regression equations that are indicated by the path diagram.

#### First the model in Fig. 7
**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

@@

**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

**Independence Set - Fig. 7**

@@

**Step 3:** For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*. The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*. The basis set can be determined by using functions in the `R` package `ggm`.

@@

**The Basis Set of Independence Claims - Fig. 7**

@@

**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*. In other words, perform a regression model using an appropriate method. Here I am only going to formulate the models because I am still in the process of calculating variables, so I do not have values for calculating probabilities of the coefficients. The predictor variable whose independence is being checked in each model is bolded - this is the variable for which the exact probability must be recorded. I also hazard an educated guess at the distribution family of the response variable for each model.

**Model Set - Fig. 7**

@@

**Step 5:** Combine the *k* probabilities into the *C*-statistic:

@@

#### The model in Fig. 8

**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

@@

**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

**Independence Sets - Fig. 8**

@@

**The Basis Set of Independence Claims - Fig. 8**

@@

**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.

**Model set - Fig. 8**

>>>>

---
title: "Appendix 2 - Fitting proposed path models using the C-statistic"
author: "Whalen Dillon"
date: "December 17, 2014"
output: 
  word_document:
    toc: true
---
----

This appendix details steps 1-4 of calculating the *C*-statistic for fitting a path model in a generalized multilevel context (Shipley 2009). I go through these steps for the path models in Fig. 7 and Fig. 8 (included here)  to produce the regression equations that are indicated by the path diagram.

# The model in Fig. 7
## Step 1
Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

@@

## Step 2:
List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

### Independence Set - Fig. 7

@@

## Step 3
For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*. The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*. The basis set can be determined by using functions in the `R` package `ggm`.

@@

### The Basis Set of Independence Claims - Fig. 7

@@

## Step 4
For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*. In other words, perform a regression model using an appropriate method. Here I am only going to formulate the models because I am still in the process of calculating variables, so I do not have values for calculating probabilities of the coefficients. The predictor variable whose independence is being checked in each model is bolded - this is the variable for which the exact probability must be recorded. I also hazard an educated guess at the distribution family of the response variable for each model.

### Model Set - Fig. 7

@@

## Step 5 
Combine the *k* probabilities into the *C*-statistic:

@@

# The model in Fig. 8

## Step 1
Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

@@

## Step 2
List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

### Independence Sets - Fig. 8

@@

### The Basis Set of Independence Claims - Fig. 8

@@

## Step 4
For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.

### Model set - Fig. 8
